Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision

### 2023-04-18 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
finish p-value

<p> 1. Generate independent random draws \(w_i, i=1,...,N\) from \(N(0, 1)\). </p>

@@

<p> 2. Regress \(w\) on \(x\) and obtain the residuals \(r_i, i=1,...,N\). </p>

@@

>>>>

<p> 1. Generate independent random draws \(w_i, i=1,...,n\) from \(N(0, 1)\). </p>

@@

<p> 2. Regress \(w\) on \(x\) and obtain the residuals \(r_i, i=1,...,n\). </p>

@@

.center[
# 📈Statistical Significance📈
]

Under the null hypothesis, it is expected that the data plot would have **no distinguishable difference** with the null plots.

Suppose observers are allowed to **select only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.

However, if we involve $K$ **independent observers** in a visual test, and show them the **same lineup**, there will be possible **dependencies** in the visual test due to **repeated evaluations** of the same lineup.

```{r}
alpha_sample <- rgamma(10, 1, 1)
data.frame(x = 0, y = 0, k = 1:20, alpha = alpha_sample/sum(alpha_sample)) %>%
ggplot() +
  geom_text(aes(x, y, label = paste0("theta[", k, "]~'='~", round(alpha, 2))), parse = TRUE, size = 5) +
  facet_wrap(~k) +
  theme_light(base_size = 12) +
  theme(panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

.center[
.caption[
Fig. 5: _Example of the probability of each plot being selected \\(\theta_i\\) under the null hypothesis with parameter \\(\alpha = 1\\)._
]
]

# 📈Statistical Significance📈

VanderPlas et al. (2021) addresses this by modelling the probability of a plot $i$ being selected from a lineup as $\theta_i$:

$$\theta_i \sim Dirichlet(\alpha) \text{ for } i=1,...,m \text{ and } \alpha > 0.$$

- This assumes the **attractiveness distribution** of each plot.

For **small** $\alpha$:
- Only **a few plots** are attractive and tend to be selected.

For **large** $\alpha$:
- The distribution of the probability of each plot being selected is more **even**.

```{r}
data.frame(k = 1:20, 
           ci = c(1, 0, 0.25, 0, 0, 0.2, 0.45, 0, 5.95, 0, 0, 0, 0, 1.7, 0, 0.2, 0, 1, 0, 0.25)) %>%
ggplot() +
  geom_text(aes(0, 0, label = paste0("c[", k, "]~'='~", ci)), parse = TRUE, size = 5) +
  facet_wrap(~k) +
  theme_light(base_size = 12) +
  theme(panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

.center[
.caption[
Fig. 5: _Example of the number of times a plot being selected \\(c_i\\). The sum of all \\(c_i\\) is 11 because the lineup is evaluated by 11 subjects._
]
]

# 📈Statistical Significance📈

The **number of times** plot $i$ being selected in $K$ evaluations is denoted as $c_i$.

- In case subject $j$ makes **multiple selections**, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the **number of plots subject $j$ selected** for $j=1,...K$. 
- This ensures $\sum_{i}c_i=K$.

.center[
# 📈Statistical Significance📈
]

Since we are only interested in the **selections of the data plot $i$**, the **marginal model** can be simplified to a **beta-binomial model**. VanderPlas et al. (2021) derives the visual $p$-value as

$$P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\text{ for } c_i \in \mathbb{Z}_0^+$$

where $B(.)$ is the beta function defined as

$$B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\text{ where } a,b>0.$$
We extend the equation to **non-negative real number $c_i$** by applying a linear approximation

$$
P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \text{ for } c_i \in \mathbb{R}_0^+.
$$
